---
title: "Smart Iris Classifier"
author: "Mark Zhang"
date: "July 21, 2018"
output:
  ioslides_presentation: default
  slidy_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
require(caret)
inTrain <- createDataPartition(y = iris$Species, p = 0.8, list = F)
trainSet <- iris[inTrain,]
testSet <- iris[-inTrain,]
```

## Introduction

The Smart Iris Classifier is capable of classfying different species of Iris given the length and width of sepal and petal. By entering the 4 measurements of an Iris, the classifier will tell you whether it's setosa, versicolor, or virginica. The test accuracy is above 95%!

## How it works?

The core engine of this classifier is a Random Forest algorithm trained on 150 rows of data. With the cutting-edge technologies such as RStudio and Caret, the process of training can be done fairly efficiently.

Here's the code:
```{r, echo=TRUE}
trControl <- trainControl(method = 'cv', number = 10)
fit_rf <- train(Species ~ ., 
                method = 'rf',
                data = trainSet,
                trControl = trControl)
```

## Test accuracy

```{r, echo = T, comment=''}
confusionMatrix(predict(fit_rf$finalModel, testSet), testSet$Species)
```

## Conclusion

- Random Forest does the magic behind the scene
- This was made possible by the *era of data*
- The combination of RStudio, Coursera, and Johns Hopkins saves the day!
